Supplementary Materials for Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs

### 2020-10-22 notes/manuscripts/AISTATS/supplement_minimax_optimal_laplacian_smoothing.tex
Final draft of AISTATS supplement.

\documentclass[twoside]{article}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}
\usepackage{fullpage}

@@

@@

\title{Supplementary Materials for Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs}
\date{}
\maketitle

In this supplement, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''. Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning. In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$. In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, and \ref{sec:empirical_norm}, we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$. It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$. In Section~\ref{sec:main_results}, we prove our main theorems and some other results. In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.

@@

To prove Lemma~\ref{lem:graph_sobolev_seminorm}, we will show that for any $f \in H^1(\Xset)$,
\begin{equation*}
\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq p_{\max}^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
\end{equation*}
whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^T \Lap_{n,r} f$ is a non-negative random variable).

Since
\begin{equation*}
f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 \mathbf{W}_{ij},
\end{equation*}
it follows that
\begin{equation}
\label{pf:first_order_graph_sobolev_seminorm_1}
\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
\end{equation}
where $X$ and $X'$ are random variables independently drawn from $P$.

For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$. We may do so without loss of generality: $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Xset)$. Obviously
\begin{equation}
\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] \leq p_{\max}^2 \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx. \label{pf:first_order_graph_sobolev_seminorm_2}
\end{equation}
and it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
\begin{align}
\int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K_r(x',x) \,dx' \,dx & = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dx' \,dx \nonumber \\
& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) \,dt\,dx' \,dx \nonumber \\
& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) \,dt\,dz \,dx \\
&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
& = r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
\end{align}
where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ and $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
\begin{equation*}
x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d.
\end{equation*}
Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
\begin{align*}
\int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{\Rd} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
& = \sigma_K \|\nabla f(\wt{x})\|^2
\end{align*}
and by plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain the claimed result.

@@

Note that the bound on $\Var[\varDelta_{i j}]$ follows as long as we can control $\|\nabla f\|_{\Leb^4(\Xset)}$; this implies the Lipschitz assumption---which gives us control of $\|\nabla f\|_{\Leb^{\infty}(\Xset)}$---can be weakened. However, the Sobolev assumption---which gives us control only over $\|\nabla f\|_{\Leb^2(\Xset)}$---will not do the job. 
\section{Bounds on neighborhood graph eigenvalues}
\label{sec:graph_eigenvalues}
In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citep{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary. The bulk of our technical effort will stem from proving Theorem~\ref{thm:neighborhood_eigenvalue}, which gives estimates on eigenvalues of the graph Laplacian in terms of eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$. We recall $\Delta_P$ is defined as
\begin{equation*}
\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x)\bigr)
\end{equation*}
To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions}. Some other notation: throughout this section, we will write $A, A_1,\ldots$ and $a,a_1,\ldots$ for constants which do not depend on $n$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
\begin{enumerate}[label=(S\arabic*)]
	\item 
	\begin{equation*}
	\theta + \wt{\delta} \leq \frac{1}{4A_1}~~\textrm{and}~~C_0\biggl(\frac{\log n}{n}\biggr)^{1/d} \leq \wt{\delta} \leq r
	\end{equation*}
\end{enumerate}

\begin{theorem}
	\label{thm:neighborhood_eigenvalue}
	For any $\ell \in \mathbb{N}$ and $\theta$ and $\wt{\delta}$ such that
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	1 - A\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + \theta + \wt{\delta}\Biggr)\geq \frac{1}{2}
	\end{equation}
	with probability at least $1 - A_1n\exp(-a_1n\theta^2\wt{\delta}^{m})$, it holds that
	\begin{equation}
	\label{eqn:eigenvalue_bound}
	a \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
	\end{equation}
\end{theorem}
We will return to proving Theorem~\ref{thm:neighborhood_eigenvalue}, but first we show that under our regularity conditions on $p$ and $\Xset$, it implies Lemma~\ref{lem:neighborhood_eigenvalue}. The link between the two is Weyl's Law.
\begin{proposition}[Weyl's Law]
	\label{prop:weyl}
	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}. Then there exist constants $a_2$ and $A_2$ such that
	\begin{equation}
	\label{eqn:weyls_law}
	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
	\end{equation}
\end{proposition}
See Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
\paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
Put
\begin{equation*}
\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
\end{equation*}
and note that the condition $r < c_0$ guarantees $\ell_{\star} \geq 2$ when $c_0$ is chosen to be a sufficiently small constant. By Proposition~\ref{prop:weyl}, we have that
\begin{align*}
\sqrt{\lambda_{\ell_{\star}}(\Delta)} \leq \frac{1}{r}\biggl(\frac{1}{2A_1} - (\theta + \wt{\delta})\biggr) 
\end{align*}
so that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied. We will therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}. Along with~\eqref{eqn:weyls_law}, this implies the following bounds on the graph Laplacian eigenvalues:
\begin{equation*}
\frac{a_2}{A_1} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{A_2}{a_1} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}.
\end{equation*}
It remains to bound $\lambda_k(G_{n,r})$ for those indices $k$ which are greater than $\ell_{\star}$. On the one hand, since the eigenvalues are sorted in ascending order, we can use the lower bound on $\lambda_{\ell_{\star}}(G_{n,r})$ that we have just derived:
\begin{align*}
\lambda_k(G_{n,r}) & \geq \lambda_{\ell_{\star}}(G_{n,r}) \\
& \geq \frac{a_2}{A_1}nr^{d + 2}\ell_{\star}^{2/d} \\
& = \frac{a_2}{A_1}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A_1) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^2 \\
& \geq \frac{a_2}{32A_1^3 A_2} nr^{d}
\end{align*}
with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A_1}$. On the other hand, for any graph $G$ the maximum eigenvalue of the Laplacian is upper bounded by twice the maximum degree \citep{chung97}. Writing $D_{\max}(G_{n,r})$ for the maximum degree of $G_{n,r}$, we show in Lemma~\ref{lem:max_degree} that
\begin{equation*}
\lambda_k(G_{n,r}) \leq 2D_{\max}(G_{n,r}) \leq 4p_{\max}nr^d 
\end{equation*}
with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$. Thus we have established each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue}.

@@

In this section we prove Theorem~\ref{thm:neighborhood_eigenvalue}, following closely the approach of \citep{burago2014,trillos2019,calder2019}. As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Dirichlet energies
\begin{equation*}
b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u 
\end{equation*}
and
\begin{equation*}
D_2(f) :=
\begin{cases*}
\int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx~~ &\textrm{if $f \in H^1(\Xset)$} \\
\infty~~ & \textrm{otherwise,}
\end{cases*}
\end{equation*}
We recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and we let the class $\Leb^2(P_n)$ consist of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$. For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is simply
\begin{equation*}
\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx,
\end{equation*} 
and we let $\Leb^2(P)$ be the set of functions for which $\|f\|_P < \infty$.

Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$. Suppose one could show the following two results: 
\begin{enumerate}[(1)]
	\item an upper bound of $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(\mc{X})$, and vice versa upper bound $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(\mc{X}) \to \Leb^2(P_n)$,
	\item that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{P} \approx \|u\|_{n}$ and $\|\mc{P}(f)\|_{P} \approx \|f\|_{n}$.
\end{enumerate}
Then, by using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e. the Courant-Fischer Theorem---one could obtain estimates on the error $\bigl|nr^{d + 2}\lambda_k(\Delta_P) - \lambda_k(G_{n,r})\bigr|$.

We will momentarily define maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$ that satisfy both (1) and (2). These maps will be defined with respect to a particular probability measure $\wt{P}_n$, that with high probability is close in transportation distance to both $P_n$ and $P$. This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate the random energy $b_r$ to the non-random $D_2$.

>>>>

\documentclass[twoside]{article}
\usepackage{aistats2021}
\usepackage{amsmath, amsfonts, amsthm, amssymb}
\usepackage[parfill]{parskip}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{verbatim}

@@

%%% New section numbering
\renewcommand{\thesection}{A\arabic{section}}

@@

\onecolumn	
\aistatstitle{Supplementary Materials for ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''}

In this supplement, we provide proofs of all results found in the paper ``Minimax Optimal Regression over Sobolev Spaces via Laplacian Regularization on Neighborhood Graphs''. Our main theorems (Theorems~\ref{thm:laplacian_smoothing_estimation1}-\ref{thm:laplacian_smoothing_testing_manifold}) all follow the same general proof strategy, centered on conditioning. In Section~\ref{sec:fixed_graph_error_bounds}, we establish (estimation or testing) error bounds which hold for any graph $G$; these bounds are stated with respect to (functionals of) the graph $G$, and allow us to upper bound the error of $\wh{f}$ and $\wh{\varphi}$ conditional on the design $\mathbf{X} = x$. In Sections~\ref{sec:graph_sobolev_seminorm}, \ref{sec:graph_eigenvalues}, \ref{sec:empirical_norm}, and \ref{sec:manifold} we give all our necessary probabilistic estimates on these functionals, for the particular random neighborhood graph $G = G_{n,r}$. It is in these sections where we invoke our various assumptions on the distribution $P$ and regression function $f_0$. In Section~\ref{sec:main_results}, we prove our main theorems and some other results. In Section~\ref{sec:concentration}, we state a few concentration bounds that we use repeatedly in our proofs.

\vfill

@@

To prove Lemma~\ref{lem:graph_sobolev_seminorm}, we will show that for any $f \in H^1(\Xset)$,
\begin{equation*}
\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] \leq (1 + L_pr)^2 \sigma_K n^2 r^{d + 2} |f|_{H^1(\Xset)}^2
\end{equation*}
whence the claim follows immediately by Markov's inequality (recall that $\Lap_{n,r}$ is positive semi-definite, and therefore $f^T \Lap_{n,r} f$ is a non-negative random variable).

Since
\begin{equation*}
f^T \Lap_{n,r} f = \frac{1}{2}\sum_{i, j = 1}^{n} \bigl(f(X_i) - f(X_j)\bigr)^2 W_{ij},
\end{equation*}
it follows that
\begin{equation}
\label{pf:first_order_graph_sobolev_seminorm_1}
\Ebb\Bigl[f^T \Lap_{n,r} f\Bigr] = \frac{n(n - 1)}{2} \Ebb\biggl[\Bigl(f(X') - f(X)\Bigr)^2 K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr]
\end{equation}
where $X$ and $X'$ are random variables independently drawn from $P$.

For the remainder of this proof, we will assume that $f \in C^{\infty}(\Xset)$. We may do so without loss of generality: $C^{\infty}(\Xset)$ is dense in $H^1(\Xset)$ and the expectation on the right hand side of~\eqref{pf:first_order_graph_sobolev_seminorm_1} is continuous in $H^1(\Xset)$. Obviously
\begin{equation}
\Ebb\biggl[\bigl(f(X') - f(X)\bigr)^2K\biggl(\frac{\|X' - X\|}{r}\biggr)\biggr] = \int_{\Xset} \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx, \label{pf:first_order_graph_sobolev_seminorm_2}
\end{equation}
so that it remains now to bound the integral. Taking a first-order Taylor expansion of $f$, we get
\begin{align}
\int_{\Xset} & \int_{\Xset} \bigl(f(x') - f(x)\bigr)^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \\
& = \int_{\Xset} \int_{\Xset} \biggl[\int_{0}^{1} \nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\,dt\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dx' \,dx \nonumber \\
& \overset{(i)}{\leq} \int_{\Xset} \int_{\Xset} \int_{0}^{1} \biggl[\nabla f\bigl(x + t(x' - x)\bigr)^T (x' - x)\biggr]^2 K\biggl(\frac{\|x' - x\|}{r}\biggr) p(x') p(x) \,dt\,dx' \,dx \nonumber \\
& \overset{(ii)}{=} r^{d + 2} \int_{\Xset} \int_{\mc{Z}_r(x)} \int_{0}^{1} \biggl[\nabla f\bigl(x + trz\bigr)^T z\biggr]^2 K\bigl(\|z\|\bigr) p(zr + x) p(x) \,dt\,dz \,dx \\
&  \overset{(iii)}{\leq} r^{d + 2} \int_{\Xset} \int_{\Reals^d} \int_{0}^{1} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p\bigl(\wt{x} + [1-t]rz\bigr) p\bigl(\wt{x} + trz\bigr) \,dt \,dz \,d\wt{x} \nonumber \\
& \overset{(iv)}\leq \biggl(1 + \frac{L_pr}{p_{\min}}\biggr)^2 r^{d + 2} \int_{\Xset} \int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) p(\wt{x})^2 \,dz \,d\wt{x} \label{pf:first_order_graph_sobolev_seminorm_3}
\end{align}
where $(i)$ follows by Jensen's inequality, $(ii)$ follows by substituting $z = (x' - x)/r$ with $\mc{Z}_r(x) := \{z: zr + x \in \Xset\}$ $(iii)$ by substituting $\wt{x} = x + trz$ and noting that
\begin{equation*}
x \in \Xset ~~\textrm{and}~~ z \in \mc{Z}_r(x) \Longrightarrow x + trz \in \Xset ~~\textrm{and}~~ z \in \Reals^d,
\end{equation*}
and $(iv)$ by the Lipschitz property of $p$.

Now, writing $\bigl(\nabla f(\wt{x}) ^T z\bigr)^2 = \bigl(\sum_{i = 1}^{d} z_{i} f^{(e_i)}(x) \bigr)^2$, expanding the square and integrating, we have that for any $\wt{x} \in \Xset$,
\begin{align*}
\int_{\Reals^d} \Bigl[\nabla f\bigl(\wt{x}\bigr)^T z\Bigr]^2 K\bigl(\|z\|\bigr) \,dz & = \sum_{i,j = 1}^{d} f^{(e_i)}(\wt{x}) f^{(e_j)}(\wt{x}) \int_{\Rd} z_iz_jK(\|z\|) \,dz \\
& = \sum_{i = 1}^{d} \bigl(f^{(e_i)}(\wt{x})\bigr)^2 \int_{\Rd} z_i^2 K\bigl(\|z\|\bigr) \,dz \\
& = \sigma_K \|\nabla f(\wt{x})\|^2
\end{align*}
and by plugging back into~\eqref{pf:first_order_graph_sobolev_seminorm_3}, we obtain the claimed result.

@@

Note that the bound on $\Var[\varDelta_{i j}]$ follows as long as we can control $\|\nabla f\|_{\Leb^4(\Xset)}$; this implies the Lipschitz assumption---which gives us control of $\|\nabla f\|_{\Leb^{\infty}(\Xset)}$---can be weakened. However, the Sobolev assumption---which gives us control only over $\|\nabla f\|_{\Leb^2(\Xset)}$---will not do the job. 
\section{Bounds on neighborhood graph eigenvalues}
\label{sec:graph_eigenvalues}
In this section, we prove Lemma~\ref{lem:neighborhood_eigenvalue}, following the lead of~\citep{burago2014,trillos2019,calder2019}, who establish similar results with respect to a manifold without boundary. The bulk of our technical effort will stem from proving Theorem~\ref{thm:neighborhood_eigenvalue}, which gives estimates on eigenvalues of the graph Laplacian in terms of eigenvalues of the weighted Laplace-Beltrami operator $\Delta_P$. We recall $\Delta_P$ is defined as
\begin{equation*}
\Delta_Pf(x) := -\frac{1}{p} \dive\bigl(p^2\nabla f(x)\bigr)
\end{equation*}
To avoid confusion, in this section we write $\lambda_k(G_{n,r})$ for the $k$th smallest eigenvalue of the graph Laplacian matrix $\Lap_{n,r}$ and $\lambda_k(\Delta_P)$ for the $k$th smallest eigenvalue of $\Delta_P$ \footnote{Under the assumptions~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}, the operator $\Delta_p$ has a discrete spectrum; see \cite{garciatrillos18} for more details.}. Some other notation: throughout this section, we will write $A, A_0, A_1,\ldots$ and $a,a_0,a_1,\ldots$ for constants which may depend on $\Xset$, $d$, $K$, and $p$ but do not depend on $n$. We let $L_K$ denote the Lipschitz constant of the kernel $K$, and we let $\theta$ and $\wt{\delta}$ denote (small) positive numbers, which always satisfy the following restrictions.
\begin{enumerate}[label=(S\arabic*)]
	\item 
	\label{asmp:smallness}
	\begin{equation*}
	\theta + \wt{\delta} \leq \frac{1}{4\max\{A_1,A\}}~~\textrm{and}~~n^{-1/d} \leq \wt{\delta} \leq \min\biggl\{\frac{1}{2^{d + 3}A_0}, \frac{1}{A_3}, \frac{K(1)}{8L_KA_0}\biggr\}r
	\end{equation*}
\end{enumerate}

\begin{theorem}
	\label{thm:neighborhood_eigenvalue}
	For any $\ell \in \mathbb{N}$ and $\theta$ and $\wt{\delta}$ such that
	\begin{equation}
	\label{eqn:neighborhood_eigenvalue_1}
	1 - A\Biggl(r \sqrt{\lambda_{\ell}(\Delta_P)} + \theta + \wt{\delta}\Biggr)\geq \frac{1}{2}
	\end{equation}
	with probability at least $1 - A_0n\exp(-a_0n\theta^2\wt{\delta}^{m})$, it holds that
	\begin{equation}
	\label{eqn:eigenvalue_bound}
	a \lambda_k(G_{n,r}) \leq nr^{d+2} \lambda_k(\Delta_P) \leq A \lambda_k(G_{n,r}),~~\textrm{for all $1 \leq k \leq \ell$}
	\end{equation}
\end{theorem}
We will return to proving Theorem~\ref{thm:neighborhood_eigenvalue}, but first we show that under our regularity conditions on $p$ and $\Xset$, it implies Lemma~\ref{lem:neighborhood_eigenvalue}. The link between the two is Weyl's Law.
\begin{proposition}[Weyl's Law]
	\label{prop:weyl}
	Suppose the density $p$ and the domain $\Xset$ satisfy~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density}. Then there exist constants $a_2$ and $A_2$ such that
	\begin{equation}
	\label{eqn:weyls_law}
	a_2k^{2/d} \leq \lambda_k(\Delta_P) \leq A_2k^{2/d}~~\textrm{for all $k \in \mathbb{N}, k > 1$}.
	\end{equation}
\end{proposition}
See Lemma 28 of~\citep{dunlop2020} for a proof that~\ref{asmp:domain} and~\ref{asmp:bounded_lipschitz_density} imply Weyl's Law.
\paragraph{Proof of Lemma~\ref{lem:neighborhood_eigenvalue}.}
Put
\begin{equation*}
\ell_{\star} = \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d},
\end{equation*}
and note that the condition $r < c_0$ guarantees $\ell_{\star} \geq 2$ when $c_0$ is chosen to be a sufficiently small constant. By Proposition~\ref{prop:weyl}, we have that
\begin{align*}
\sqrt{\lambda_{\ell_{\star}}(\Delta_P)} \leq \frac{1}{r}\biggl(\frac{1}{2A} - (\theta + \wt{\delta})\biggr) 
\end{align*}
so that condition~\eqref{eqn:neighborhood_eigenvalue_1} is satisfied. We may therefore assume the conclusions of Theorem~\ref{thm:neighborhood_eigenvalue}. Along with~\eqref{eqn:weyls_law}, this implies the following bounds on the graph Laplacian eigenvalues:
\begin{equation*}
\frac{a}{A_2} nr^{d + 2} k^{2/d} \leq \lambda_k(G_{n,r}) \leq \frac{A}{a_2} nr^{d + 2} k^{2/d}~~\textrm{for all $2 \leq k \leq \ell_{\star}$}.
\end{equation*}
It remains to bound $\lambda_k(G_{n,r})$ for those indices $k$ which are greater than $\ell_{\star}$. On the one hand, since the eigenvalues are sorted in ascending order, we can use the lower bound on $\lambda_{\ell_{\star}}(G_{n,r})$ that we have just derived:
\begin{align*}
\lambda_k(G_{n,r}) & \geq \lambda_{\ell_{\star}}(G_{n,r}) \\
& \geq \frac{a_2}{A}nr^{d + 2}\ell_{\star}^{2/d} \\
& = \frac{a_2}{A}nr^{d + 2} \floor{\biggl(\frac{\bigl(1/(2A) - (\theta + \wt{\delta})\bigr)}{rA_2^{1/2}}\biggr)^d}^{2/d} \\
& \geq \frac{a_2}{32A^3 A_2} nr^{d}
\end{align*}
with the last line following since $\theta + \wt{\delta} \leq \frac{1}{4A}$. On the other hand, for any graph $G$ the maximum eigenvalue of the Laplacian is upper bounded by twice the maximum degree \citep{chung97}. Writing $D_{\max}(G_{n,r})$ for the maximum degree of $G_{n,r}$, we show in Lemma~\ref{lem:max_degree} that
\begin{equation*}
\lambda_k(G_{n,r}) \leq 2D_{\max}(G_{n,r}) \leq 4p_{\max}nr^d 
\end{equation*}
with probability at least $1 - 2n\exp\Bigl(-nr^dp_{\min}/(3K(0)^2)\Bigr)$. Setting $\theta$ and $\wt{\delta}$ as large as possible while ensuring that~\ref{asmp:smallness} and~\eqref{eqn:neighborhood_eigenvalue_1} are satisfied---so that $\theta \asymp 1$ and $\wt{\delta} \asymp r$---we have established that each of the inequalities in Lemma~\ref{lem:neighborhood_eigenvalue} hold with probability at least $1 - Cn\exp(-cnr^d)$.

@@

In this section we prove Theorem~\ref{thm:neighborhood_eigenvalue}, following closely the approach of \citep{burago2014,trillos2019,calder2019}. As in these works, we relate $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$ by means of the Dirichlet energies
\begin{equation*}
b_r(u) := \frac{1}{n^2 r^{d+ 2}}u^T \Lap_{n,r} u 
\end{equation*}
and
\begin{equation*}
D_2(f) :=
\begin{cases*}
\int_{\Xset} \|\nabla f(x)\|^2 p^2(x) \,dx~~ &\textrm{if $f \in H^1(\Xset)$} \\
\infty~~ & \textrm{otherwise,}
\end{cases*}
\end{equation*}
We recall that for a function $u: \{X_1,\ldots,X_n\} \to \Reals$, the empirical norm is defined as $\|u\|_n^2 := \frac{1}{n} \sum_{i = 1}^{n} (u(X_i))^2$, and we let the class $\Leb^2(P_n)$ consist of those $u \in \Reals^n$ for which $\|u\|_{n} < \infty$. For a function $f: \Xset \to \Reals$, the continuum limit of $\|f\|_n^2$ is 
\begin{equation*}
\|f\|_{P}^2 := \int_{\Xset} \bigl|f(x)\bigr|^2 p(x) \,dx.
\end{equation*}

Let us pause briefly to motivate the relevance of $b_r(u)$ and $D_2(f)$. Suppose one could show the following two results: 
\begin{enumerate}[(1)]
	\item an upper bound of $b_r(u)$ by $D_2\bigl(\mc{I}(u)\bigr)$ for an appropriate choice of interpolating map $\mc{I}: \Leb^2(P_n) \to \Leb^2(\mc{X})$, and vice versa an upper bound of $D_2(f)$ by $b_r(\mc{P}(f))$ for an appropriate choice of discretization map $\mc{P}: \Leb^2(\mc{X}) \to \Leb^2(P_n)$,
	\item that $\mc{I}$ and $\mc{P}$ were near-isometries, meaning $\|\mc{I}(u)\|_{P} \approx \|u\|_{n}$ and $\|\mc{P}(f)\|_{P} \approx \|f\|_{n}$.
\end{enumerate}
Then, by using the variational characterization of eigenvalues $\lambda_k(\Delta_P)$ and $\lambda_k(G_{n,r})$---i.e. the Courant-Fischer Theorem---one could obtain estimates on the error $\bigl|nr^{d + 2}\lambda_k(\Delta_P) - \lambda_k(G_{n,r})\bigr|$.

We will momentarily define particular maps $\wt{\mc{I}}$ and $\wt{\mc{P}}$, and establish that they satisfy both (1) and (2). In order to define these maps, we must first introduce a particular probability measure $\wt{P}_n$ that with high probability is close in transportation distance to both $P_n$ and $P$. This estimate on the transportation distance---which we now give--- will be the workhorse that allows us to relate $b_r$ to $D_2$, and $\|\cdot\|_n$ to $\|\cdot\|_P$.

\paragraph{Transportation distance between $P_n$ and $P$.}
For a measure $\mu$ defined on $\Xset$ and map $T: \Xset \to \Xset$, let $T_{\sharp}\mu$ denote the \emph{push-forward} of $\mu$ by $T$, i.e the measure for which
\begin{equation*}
\bigl(T_{\sharp}\mu\bigr)(U) := \mu\bigl(T^{-1}(U)\bigr)
\end{equation*}
for any Borel subset $U \subseteq \Xset$. Suppose $T_{\sharp}\mu = P_n$; then the map $T$ is referred to as transportation map between $\mu$ and $P_n$. The  $\infty$-transportation distance between $\mu$ and $P_n$ is then
\begin{equation}
\label{eqn:optimal_transport}
d_{\infty}(\mu,P_n) := \inf_{T: T_{\sharp} \mu = P_n} \|T - \mathrm{Id}\|_{L^{\infty}(\mu)}
\end{equation}
where $\mathrm{Id}(x) = x$ is the identity mapping.
